---
title: "General Social Survey"
author: "Daniel Martinez"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"
mydata5 <- filter(mydata,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE")
```


# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey.The first blocks of R-code has selected deown a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data.

How are the "voted in 2016" and "voted in 2012" related?
$H_0$ voting in 2012 and 2016 are unrelated.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

use categorical variables, and a chi-squared test to test the hypothesis that voting in 2016 is / is not independent of voting in 2012

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the voting trends between 2012 and 2016

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent", xlab='voted in 2016') 

```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

numerical summary is a cross tablulation:

```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```
The hypothesis was that voting in 2012 related to voting in 2016.We have evidence that if you voted in 2012, you were likely to vote in 2016. We also have evidence that if you did not vote in 2012, you were likely to not vote in 2016.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

We examine if there is a relationship in the belief in the bigbang theory (BIGBANG) with the belief in the afterlife (POSTLIFE)

$H_0$ belief in the bigbang theory is unrelated to belief in the afterlife.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Use categorical variables, and a chi-squared test to test the hypothesis that belief in the bigbang theory is independent to belief in the afterlife

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
we use a barchart to show the percentage of belief in Bigbangtheory and belief in the afterlife among surveyed voters

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent",'Believe in Postlife',main='BELIEVE IN BIGBANG') 

```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical summary is a cross tabulation
```{r}
xtabs(~BIGBANG+POSTLIFE, data=mydata3)
```
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->



```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
Bangtab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(Bangtab)
```

We reject the null hypothesis.There is some relationship between belief in the Big Bang Theory and the Afterlife.

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

We compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).We are seeking to find if there is a relationship between views on death penalty (CAPPUN) or opinion on courts (COURTS) and voting likelihood in 2012 and 2016.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Use categorical variables, and a chi-squared test to test the hypothesis that likelihood of voting in 2016 and 2012 is / is not independent of opinion on death penalty or opinion on courts.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

we use a barchart to show the percentage of likelihood of voting in 2016 and 2012 with opinion on Death penalty or courts among surveyed voters


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent", xlab='voted in 2016')

barchartGC(~CAPPUN+COURTS, data=mydata4, type = "percent",'Views on Death Penalty',main='Views on Courts') 

```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
Numerical summary is a cross tabulation
```{r}
xtabs(~CAPPUN+COURTS, data=mydata4)
```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata4)
viewtab=xtabs(~CAPPUN+COURTS, data=mydata4)
fisher.test(viewtab)
```
There appears to be a noticable likelihood that those that favor the death penalty are more likely to view the courts and not harsh enough, and that those that oppose the death penalty are more likely to view the courts as too harsh.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The data appears to indicate that there is some relationship between past and present voting likelihood, and that the various political opinions and views have a noticable effect of the likelihood of voting.

